Category-Level Transfer Learning from Knowledge Base to Microblog Stream for Accurate Event Detection

### 2017-03-25 paper/presentation/DASFAA_presentation.tex
presentation 0.9

@@

%------------------------------
%page 4
\begin{frame}
\frametitle{Challenges (2/2)}	
Exploratory study on the \textit{Edinburgh twitter corpus}: 11/29 events contain less than 50 tweets.

@@

%------------------------------
%page 8
\begin{frame}
\frametitle{Overview of our solution}

\begin{figure}[h]
		\setlength{\abovecaptionskip}{0.cm}
        \setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=0.5\columnwidth]{../img/NSDetectorExample.pdf}
        \caption{\textsc{TransDetector}'s processing flow, taking \textit{Military} related events in microblogs as an example.}
        \label{fig:hood}
\end{figure}

@@

\begin{comment}
%------------------------------
%page 9
\begin{frame}
\frametitle{\textsc{TransDetector}: Phrase 1}	
\begin{figure}[h]
		\setlength{\abovecaptionskip}{0.cm}
        \setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=0.5\columnwidth]{../img/initializationExample.pdf}
        \caption{Extracting Category-Level Topics in Knowledge Base via its three fold hierarchical structure, taking \textit{Military} as an example.}
        \label{fig:hood}
\end{figure}

@@

%------------------------------
%page 10
\begin{frame}
\frametitle{\textsc{TransDetector}: Phrase 2}

@@

%------------------------------
%page 11
\begin{frame}
\frametitle{\textsc{TransDetector}: Phrase 3}

@@

%------------------------------
%page 12
\begin{frame}
\frametitle{Experimental Results}	
\begin{figure}[h]
	\setlength{\abovecaptionskip}{0.cm}
	\setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=1.0\columnwidth]{../img/NPMI.pdf}
        \caption{More topics are compared at the NPMI metrics between our method and LightLDA}
        \label{fig:NPMI}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{some}

@@

\begin{frame}
\frametitle{Experiment Settings}

\begin{frame}
\frametitle{Experimental Results}	
\begin{table}
\setlength{\abovecaptionskip}{0.cm}%set the distance between caption and table to 0 cm.
\setlength{\belowcaptionskip}{0.cm}
\centering
\caption{Category-Level Topics extracted from knowledge base and the corresponding topics on microblog stream learned from CTrans-LDA. The words in \textbf{\textit{bold}} font are newly learned on the microblog stream by the transfer learning.}
\scalebox{0.55}{
\begin{tabular}{|cc|cc|cc|cc|cc|cc|}
\hline
\multicolumn{2}{|c|}{\textit{Aviation}} & \multicolumn{2}{c|}{\textit{Health}} & \multicolumn{2}{c|}{\textit{Middle East}} & \multicolumn{2}{c|}{\textit{Military}} & \multicolumn{2}{c|}{\textit{Mobile Phones}}\\
\begin{tabular}[c]{@{}c@{}}Knowledge\\ Base\end{tabular} & \begin{tabular}[c]{@{}c@{}}Microblog\\ Stream\end{tabular} & \begin{tabular}[c]{@{}c@{}}Knowledge\\ Base\end{tabular} & \begin{tabular}[c]{@{}c@{}}Microblog\\ Stream\end{tabular} & \begin{tabular}[c]{@{}c@{}}Knowledge\\ Base\end{tabular} & \begin{tabular}[c]{@{}c@{}}Microblog\\ Stream\end{tabular} & \begin{tabular}[c]{@{}c@{}}Knowledge\\ Base\end{tabular} & \begin{tabular}[c]{@{}c@{}}Microblog\\ Stream\end{tabular} & \begin{tabular}[c]{@{}c@{}}Knowledge\\ Base\end{tabular} & \begin{tabular}[c]{@{}c@{}}Microblog\\ Stream\end{tabular} \\ 
\hline
aircraft & air & health & weight & al & \textbf{\textit{\#syria}} & army & killed & android & iphone\\ 
air & plane & patients & loss & israel & \textbf{\textit{\#bahrain}} & military & news & mobile & apple \\ 
airport & flight & medical & diet & iran & people & air & \textbf{\textit{\#libya}} & nokia & android \\ 
flight & time & disease & health & arab & israel & command & libya & ios & app \\
airline & airlines & treatment & cancer & israeli & police & force & rebels & phone & ipad \\
airlines & news & hospital & lose & egypt & \textbf{\textit{\#libya}} & regiment & people & samsung & samsung \\
aviation & boat & patient & fat & egyptian & \#egypt & forces & police & game & mobile\\
flying & airport & clinical & tips & ibn & news & squadron & war & app & blackberry \\
pilot & force & symptoms & treatment & jerusalem & \textbf{\textit{\#israel}} & infantry & libyan & iphone & tablet \\
squadron & fly & cancer & body & syria & world & battle & attack & htc & apps\\
\hline
\end{tabular}
}
\label{tbl:historyStates}
\end{table}
\end{frame}

@@

@@

\end{comment}

>>>>

%used for algorithm
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{algorithmic}

@@

%------------------------------
%page 4
\begin{frame}
\frametitle{Challenges (2/2)}	
Exploratory study on the \textit{Edinburgh twitter corpus}: 11/27 events contain less than 50 tweets.

@@

%------------------------------
%page 8
\begin{frame}
\frametitle{Our solution}	
\textsc{TransDetector}: a novel category-level transfer learning method
\begin{itemize}
	\item transfer KB's \textbf{category-level info} into microblog stream
	\item balance the performance and the cost of leveraging knowledge base for event detection
\end{itemize}

\intextsep=5pt plus 3pt minus 1pt % 调整表格与正文的上间距
\begin{figure}[h]
		\setlength{\abovecaptionskip}{0.cm}
        \setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=0.43\columnwidth]{../img/NSDetectorExample.pdf}
        \caption{\textsc{TransDetector}'s processing flow, in 3 phases.}
        \label{fig:hood}
\end{figure}

@@

\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 1 (Extracting Category-Level Topics in KB)}	
There is a three fold hierarchical structure in Knowledge Base.
\begin{enumerate}
	\item \textbf{Taxonomy Graph \(G^{(0)}\)}. The directed edges in \(G^{(0)}\) represent the \textit{class}\(\rightarrow\)\textit{subclass} relations in KB. 
		\begin{itemize}
			\item e.g., \textit{Main topic classifications} \(\rightarrow\) \textit{Society}
		\end{itemize}
	\item \textbf{Category-Page Bipartite Graph \(G^{(1)}\)}. The directed edges in \(G^{(1)}\) represent the \textit{class}\(\rightarrow\)\textit{instance} relations in KB.
		\begin{itemize}
			\item e.g., \textit{Military} \(\rightarrow\) \textit{page: Armed Forces}
		\end{itemize}
	\item \textbf{Page-Content Map \(G^{(2)}\)}. 
For a specific Wikipedia dumps version, the edges  \textit{page} \(\rightarrow\)\textit{content} in \(G^{(2)}\) define a one-to-one mapping.
		\begin{itemize}
			\item e.g., \textit{page: Armed Forces} \(\rightarrow\) \textit{content(20170325version): ``The armed forces of a country are its government-sponsored defense, fighting forces, and organizations..."}
		\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 1 (Extracting Category-Level Topics in KB)}
\begin{algorithm}[H]
\scriptsize
\caption{Extraction of Category-Level Topics in Knowledge Base}
\label{alg:normalStatesInit}

\KwIn{Taxonomy's Graph \(G^{(0)}\), Category-Page Bipartite Graph \(G^{(1)}\), Page-Content Bipartite Graph \(G^{(2)}\), topic related category node \(c\)}
\KwOut{\(c\)'s category-level topic in knowledge base \(\bm{h}_c\)}
\(Pages(c)\leftarrow \varnothing\), \(\bm{h}_c \leftarrow \varnothing\)\\
DAG \(G^{(0)'} \leftarrow\) Remove Cycles of \(G^{(0)}\) by nodes' HITS-PageRank scores. \label{alg:line2inNormalStatesInit}\\
\(SuccessorNodes(c) \leftarrow \) Breadth-first-traverse(\(G^{(0)'},c\))\label{alg:line3inNormalStatesInit}\\
\For{\(node \in SuccessorNodes(c)\)}{
    \(Pages(c) \leftarrow Pages(c) \cup G^{(1)}.neighbours(node)\) \label{alg:line5inNormalStatesInit}\\
}
Word frequency table \(n(c,.) \leftarrow \) do word count on the text contents of \(Pages(c)\) \\
Word frequency table \(n(All,.) \leftarrow \) do word count on the text contents of all pages in \(G^{(2)}\).\\
\For{word \(w\) in WordFrequencyTable(All).keys()}{
    \(chi(c,w) \leftarrow \) \(w\)'s chi-square statistics on \(WordFrequencyTable(c)\) and \(WordFrequencyTable(All)\).\\
    \(h_{c,w} \leftarrow chi(c,w)\) \label{alg:line10inNormalStatesInit}\\
}
\Return{\(\bm{h}_c\)}
\end{algorithm}

@@

%\begin{comment}
%------------------------------
%page 9
\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 1 (Extracting Category-Level Topics in KB) (1/3)}	
Taking the category \textit{Military} as an example, we extract \textit{Military}'s category-level topic \(\bm{h_{Military}}\).
\begin{figure}[h]
		\setlength{\abovecaptionskip}{0.cm}
        \setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=0.45\columnwidth]{../img/initializationExample.pdf}
        \caption{Extracting Category-Level Topics in Knowledge Base via its three fold hierarchical structure, taking \textit{Military} as an example.}
        \label{fig:hood}
\end{figure}

@@

%------------------------------
%page 10
\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 2 (Transferring Category-Level Info into Microblog Stream) (2/3)}	
Transfer KB's Category-Level Topics \(\{\bm{h_c}\}_{c=1}^{K_{KB}}\) into microblogs stream: CTrans-LDA.
\begin{figure}[h]
		\setlength{\abovecaptionskip}{0.cm}
        \setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=0.6\columnwidth]{../model/lda_tikz.pdf}
        \caption{Diagram of CTrans-LDA.}
        \label{fig:hood}
\end{figure}

In CTrans-LDA, \(\{\bm{h_c}\}_{c=1}^{K_{KB}}\) is used as prior information:
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{0pt}
\begin{scriptsize}
\begin{equation}
\label{eq:wikiPrior}
\begin{aligned}
\tau_{kv}=
\left\{ \begin{aligned}
\lambda \frac{h_{kv}}{\sum_{v\in S_{k}}h_{kv}} &,v\in S_{k}\ and  \ k \leq K_{\bm{KB}} \\
0&,v \notin S_{k} \ or \ k > K_{\bm{KB}} \\
\end{aligned}\right.
\end{aligned}
\end{equation}
\end{scriptsize}
\end{frame}

\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 2 (Transferring Category-Level Info into Microblog Stream) (3/3)}	
We use Gibbs Sampling for solving CTrans-LDA.
\begin{itemize}
	\item The initialization probability \(\hat{q}_{k|v}\) makes sure that the learned topics are aligned to the pre-defined category-level topic.
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{0pt}
\begin{scriptsize} 
\begin{equation}
\label{eq:initProbability}
\begin{aligned}
\hat{q}_{k|v}=
\left\{ \begin{aligned}
\frac{\tau_{kv}}{\sum_{k=1}^{K}\tau_{kv}} &,\sum_{k}\tau_{kv}>0 & (a)\\
0&, \sum_{k}\tau_{kv}=0 \ and \ k \leq K_{\bm{KB}} & (b)\\
1/(K-K_{\bm{KB}})&,\sum_{k}\tau_{kv}=0 \ and \ k > K_{\bm{KB}} & (c)
\end{aligned}\right.
\end{aligned}
\end{equation}
\end{scriptsize}
\item Conditional probability in gibbs sampling:
\(p(z_{dn}=k|.)\propto (n^{(d)}_{dk}+\alpha m_k)(n^{(w)}_{kv}+\tau_{kv}+\beta)/(n^{(w)}_{k,.}+\tau_{k,.}+V\beta)\).
\end{itemize}

@@

%------------------------------
%page 11
\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 3 (Detecting Events on Category-Level Word Time Series) (1/2)}	
After transfer learning, we conduct analysis on category-level word time series, and detect events in microblog stream.
\begin{figure}[h]
		\setlength{\abovecaptionskip}{0.cm}
        \setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=.99\columnwidth]{../img/screenShot.png}
        \caption{Visualizing Category-Level Word Time Series in Microblog Stream on Edinburgh Twitter Corpus (20110711-20110915), taking \textit{Military} as an example.}
        \label{fig:hood}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 3 (Detecting Events on Category-Level Word Time Series) (2/2)}
With richer semantics, and fewer noise, we detect events more accurately, in all the following granularities.(see more technique details in our paper)
\begin{enumerate}
	\item events' candidate words
	\begin{itemize}
		\item e.g., \textit{Ft.}, \textit{Hood}, \textit{attack}.
	\end{itemize}
	\item event phrases
	\begin{itemize}
		\item e.g., \textit{Ft. Hood attack}.
	\end{itemize}
	\item events' representative microblogs
	\begin{itemize}
		\item e.g., \textit{Possible Ft. Hood Attack Thwarted \url{http://t.co/BSJ33hk}}.
	\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Experiment Settings}
\textbf{Dataset}
\begin{itemize}
	\item Knowledge Base. Wikipedia dumps\footnote{\url{https://dumps.wikimedia.org/enwiki/latest/enwiki-latest-categorylinks.sql.gz}}\footnote{\url{https://dumps.wikimedia.org/enwiki/latest/enwiki-latest-pages-articles.xml.bz2}}
	\item Microblog Stream. Edinburgh twitter corpus\footnote{\url{http://demeter.inf.ed.ac.uk/cross/docs/fsd_corpus.tar.gz}}
\end{itemize}

\textbf{Baseline Methods}
\begin{itemize}
	\item Twevent, BurstyBTM, LSH, EDCoW, and TimeUserLDA
\end{itemize}
\textbf{Ground Truth}
\begin{itemize}
	\item Benchmark1 is labeled events in previous study.
	\item Benchmark2 is our manually checked events based on Twevent, BurstyBTM, LSH, EDCoW, TimeUserLDA and \textsc{TransDetector}.
\end{itemize}
\end{frame}

%------------------------------
%page 12
\begin{frame}
\frametitle{Experimental Results}
Evaluation on Categroy-Level Topics in Knowledge Base. On \textit{Aviation} topic, semantic coherence is much better than LightLDA (same as LDA) in terms of NPMI[Röder, WSDM 2015].

@@

\begin{frame}
\frametitle{Experimental Results}	
Evaluation on Categroy-Level Topics in Knowledge Base.
On more topics, semantic coherence is much better than LightLDA (same as LDA) in terms of NPMI[Röder, WSDM 2015]. 
\begin{figure}[h]
	\setlength{\abovecaptionskip}{0.cm}
	\setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=1.0\columnwidth]{../img/NPMI.pdf}
        \caption{More topics are compared at the NPMI metrics between our method and LightLDA}
        \label{fig:NPMI}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Experimental Results}	
Effectiveness of transferring category-level topics into the microblog stream, and finding more new words in the stream which is not stored in the knowledge base.

\begin{table}
\setlength{\abovecaptionskip}{0.cm}%set the distance between caption and table to 0 cm.
\setlength{\belowcaptionskip}{0.cm}
\centering
\caption{Category-Level Topics extracted from knowledge base and the corresponding topics on microblog stream learned from CTrans-LDA. The words in \textbf{\textit{bold}} font are newly learned on the microblog stream by the transfer learning.}
\scalebox{0.55}{
\begin{tabular}{|cc|cc|cc|cc|cc|cc|}
\hline
\multicolumn{2}{|c|}{\textit{Aviation}} & \multicolumn{2}{c|}{\textit{Health}} & \multicolumn{2}{c|}{\textit{Middle East}} & \multicolumn{2}{c|}{\textit{Military}} & \multicolumn{2}{c|}{\textit{Mobile Phones}}\\
\begin{tabular}[c]{@{}c@{}}Knowledge\\ Base\end{tabular} & \begin{tabular}[c]{@{}c@{}}Microblog\\ Stream\end{tabular} & \begin{tabular}[c]{@{}c@{}}Knowledge\\ Base\end{tabular} & \begin{tabular}[c]{@{}c@{}}Microblog\\ Stream\end{tabular} & \begin{tabular}[c]{@{}c@{}}Knowledge\\ Base\end{tabular} & \begin{tabular}[c]{@{}c@{}}Microblog\\ Stream\end{tabular} & \begin{tabular}[c]{@{}c@{}}Knowledge\\ Base\end{tabular} & \begin{tabular}[c]{@{}c@{}}Microblog\\ Stream\end{tabular} & \begin{tabular}[c]{@{}c@{}}Knowledge\\ Base\end{tabular} & \begin{tabular}[c]{@{}c@{}}Microblog\\ Stream\end{tabular} \\ 
\hline
aircraft & air & health & weight & al & \textbf{\textit{\#syria}} & army & killed & android & iphone\\ 
air & plane & patients & loss & israel & \textbf{\textit{\#bahrain}} & military & news & mobile & apple \\ 
airport & flight & medical & diet & iran & people & air & \textbf{\textit{\#libya}} & nokia & android \\ 
flight & time & disease & health & arab & israel & command & libya & ios & app \\
airline & airlines & treatment & cancer & israeli & police & force & rebels & phone & ipad \\
airlines & news & hospital & lose & egypt & \textbf{\textit{\#libya}} & regiment & people & samsung & samsung \\
aviation & boat & patient & fat & egyptian & \#egypt & forces & police & game & mobile\\
flying & airport & clinical & tips & ibn & news & squadron & war & app & blackberry \\
pilot & force & symptoms & treatment & jerusalem & \textbf{\textit{\#israel}} & infantry & libyan & iphone & tablet \\
squadron & fly & cancer & body & syria & world & battle & attack & htc & apps\\
\hline
\end{tabular}
}
\label{tbl:historyStates}
\end{table}
\end{frame}

@@

\begin{frame}
\frametitle{Conclusions}	
\begin{itemize}
	\item Knowledge base is constructed elaborately and contains rich information, which can benefit the not-well-organized microblog stream. 
	\item We propose \textsc{TransDetector} method, and
	\begin{itemize}
		\item use category-level topic in knowledge base as the prior knowledge,
		\item transfer abundant knowledge from knowledge base into microblog stream
		\item enrich the semantics of microblogs and further enhances the accuracy of microblogs event detection
	\end{itemize}
\end{itemize}
\end{frame}

@@

%\end{comment}
